### 2016-11-28 conclusions.tex
section on sub-optimal network pruning

\section{Conclusions \& Future Work}
In conclusion, we must first re-assert that we do not present this work as a bench-marking study of the algorithm we derived and tested. We have merely used this algorithm as a jumping off point to investigate the nature of learning representations in neural networks. What we discovered is that first and second order methods do not make particularly good pruning decisions, and that a brute-force algorithm does surprisingly well, despite being computationally expensive. A simplistic brute-force method does so well in fact, we argue that further investigation is required to make this algorithm computationally tractable. We also observed strong evidence for the hypotheses of \cite{mozer1989skeletonization} regarding the dualist nature of neurons, i.e. that learning representations are divided between units which either participate in the output approximation or learn to cancel each others influence. This suggests that neural networks may in fact learn a minimal network implicitly, though we cannot say for sure that this is the case without further investigation. A necessary experiment in this vein would be to compare the size of network constructed using cascade correlation and compare it to the results described herein.

We have presented a novel algorithm for pruning whole neurons from a trained neural network using a second-order Taylor series approximation of the change in error resulting from the removal a given neuron as a pruning criteria. We compared this method to a first order method and a brute-force serial removal method which exhaustively found the next best single neuron to remove at each stage. Our algorithm relies on a combination of assumptions similar to the ones made by \cite{mozer1989skeletonization} and \cite{lecun1989optimal} in the formulation of the Skeletonization and Optimal Brain Damage algorithms. First, we assumed that the error function with respect to each individual neuron can be approximated with a straight line or more precisely with a parabola. Second, for second derivative terms we consider only the diagonal elements of the Hessian matrix, i.e. we assume that each neuron-weight connection can be treated independently of the other elements in the network. Third, we assumed that pruning could be done in a serial fashion in which we find the single least productive element in the network, remove it, and move on. We found that all of these assumptions are deeply flawed in the sense that the true relevance of a neuron can only be partially approximated by a first or second order method, and only at certain stages of the pruning process. For most problems, these methods can usually remove between 10-30\% of the neurons in a trained network, but beyond this point their reliability breaks down. For certain problems, none of the described methods seem to perform very well, though for obvious reasons the brute-force method always exhibits the best results. The reason for this is that the error function with respect to each hidden unit is more complex than a simple second-order Taylor series can approximate. Furthermore, we have not directly taken into account the interdependence of elements within a network, though the work of \cite{hassibi1993second} could provide some guidance in this regard. This is another critical issue to investigate in the future.

>>>>

\section{Conclusions \& Future Work}
In conclusion, we must first re-assert that we do not present this work as a bench-marking study of the algorithm we derived and tested. We have merely used this algorithm as a jumping off point to investigate the nature of learning representations in neural networks. What we discovered is that first and second order methods do not make particularly good pruning decisions, and can get hopelessly lost after making a bad pruning decision resulting in a network fault. Furthermore, the brute-force algorithm does surprisingly well, despite being computationally expensive. This method does \textit{so} well in fact, we argue that further investigation is warranted to make this algorithm computationally tractable, though we do not speculate on how that should be done here.

We also observed strong evidence for the hypotheses of \cite{mozer1989skeletonization} regarding the ``dualist'' nature of hidden units, i.e. that learning representations are divided between units which either participate in the output approximation or learn to cancel each others influence. This suggests that neural networks may in fact learn a minimal network implicitly, though we cannot say for sure that this is the case without further investigation. A necessary experiment to this end would be to compare the size of network constructed using cascade correlation (\cite{fahlman1989cascade}) and compare it to the results described herein.

We have presented a novel algorithm for pruning whole neurons from a trained neural network using a second-order Taylor series approximation of the change in error resulting from the removal a given neuron as a pruning criteria. We compared this method to a first order method and a brute-force serial removal method which exhaustively found the next best single neuron to remove at each stage. Our algorithm relies on a combination of assumptions similar to the ones made by \cite{mozer1989skeletonization} and \cite{lecun1989optimal} in the formulation of the Skeletonization and Optimal Brain Damage algorithms.

First, we assumed that the error function with respect to each individual neuron can be approximated with a straight line or more precisely with a parabola. Second, for second derivative terms we consider only the diagonal elements of the Hessian matrix, i.e. we assume that each neuron-weight connection can be treated independently of the other elements in the network. Third, we assumed that pruning could be done in a serial fashion in which we find the single least productive element in the network, remove it, and move on. We found that all of these assumptions are deeply flawed in the sense that the true relevance of a neuron can only be partially approximated by a first or second order method, and only at certain stages of the pruning process.

For most problems, these methods can usually remove between 10-30\% of the neurons in a trained network, but beyond this point their reliability breaks down. For certain problems, none of the described methods seem to perform very well, though for obvious reasons the brute-force method always exhibits the best results. The reason for this is that the error function with respect to each hidden unit is more complex than a simple second-order Taylor series can approximate. Furthermore, we have not directly taken into account the interdependence of elements within a network, though the work of \cite{hassibi1993second} could provide some guidance in this regard. This is another critical issue to investigate in the future.

### 2016-11-28 introduction.tex
finally
section on sub-optimal network pruning

Essentially, \cite{mozer1989using} observed that neural networks during training do \textit{not} distribute the learning representation evenly or equitably across hidden units. What actually happens is that a few, elite neurons learn an approximation of the input-output function, and the remaining units must learn a complex interdependence function which cancels out their respective influence on the network output. Furthermore, assuming enough units exist to learn the function in question, increasing the number of parameters does not increase the richness or robustness of the learned approximation, but rather simply increases the likelihood of overfitting and the number of noisy parameters to be canceled during training. This is evinced by the fact that in many cases, multiple neurons can be removed from a network with no re-training and with negligible impact on the quality of the output approximation. In other words, there are few bipartisan units in a trained network. A unit is typically either part of the (possibly overfit) input-output function approximation, or it is part of an elaborate noise cancellation task force. Assuming this is the case, most of the compute-time spent training a neural network is likely occupied by this arguably wasteful procedure of silencing superfluous parameters, and pruning can be viewed as a necessary procedure to ``trim the fat.''

@@

In terms of removing units without re-training, what we discovered is that predicting the behavior of a network when a unit is to be pruned is very difficult, and most of the approximation techniques put forth in existing pruning algorithms do not fare well at all when compared to a brute-force search. To begin our discussion of how we arrived at our algorithm and set up our experiments, we begin with a review of the existing literature.

>>>>

Essentially, \cite{mozer1989using} observed that during training neural networks do \textit{not} distribute the learning representation evenly or equitably across hidden units. What actually happens is that a few, elite neurons learn an approximation of the input-output function, and the remaining units must learn a complex interdependence function which cancels out their respective influence on the network output. Furthermore, assuming enough units exist to learn the function in question, increasing the number of parameters does not increase the richness or robustness of the learned approximation, but rather simply increases the likelihood of overfitting and the number of noisy parameters to be canceled during training. This is evinced by the fact that in many cases, multiple neurons can be removed from a network with no re-training and with negligible impact on the quality of the output approximation. In other words, there are few bipartisan units in a trained network. A unit is typically either part of the (possibly overfit) input-output function approximation, or it is part of an elaborate noise cancellation task force. Assuming this is the case, most of the compute-time spent training a neural network is likely occupied by this arguably wasteful procedure of silencing superfluous parameters, and pruning can be viewed as a necessary procedure to ``trim the fat.''

@@

In terms of removing units without re-training, what we discovered is that predicting the behavior of a network when a unit is to be pruned is very difficult, and most of the approximation techniques put forth in existing pruning algorithms do not fare well at all when compared to a brute-force search. To begin our discussion of how we arrived at our algorithm and set up our experiments, we review of the existing literature.

### 2016-11-28 methodology.tex
section on sub-optimal network pruning

\section{Pruning Neurons to Shrink Neural Networks}\label{sec2}
As discussed in Section \ref{sec1} our aim here is to leverage the non-uniform distribution of the learning representation in pre-trained neural networks to eliminate excess neurons in their entirety, without focusing on individual weight parameters. Taking this approach enables us to remove all the weights (incoming and outgoing) associated with a non-contributing neuron in one go. We would like to note here that in an ideal scenario, based on the neuron interdependency theory put forward by \cite{mozer1989skeletonization}, one would evaluate all possible combinations of neurons to remove (one at a time, two at a time, three at a time and so forth) to find the optimal subset of neurons to keep. This however is not feasible practically, which is why we will only focus on removing one neuron at a time and explore "greedy" algorithms to do this in an efficient manner.

The general approach taken to prune an optimally trained neural network here is to create a ranked list of all the neurons in the network based off of one of the 3 proposed ranking criteria: Brute Force approximation (which we use as our ground truth), linear approximation and quadratic approximation of the neuron's impact on the overall performance of the network. We then test the effects of removing neurons on the accuracy and error of the network. All the algorithms and methods presented here are easily parallelizable.

One last thing to note here before moving forward is that the methods discussed in this section involve some non-trivial mathematics, discussing all of which is beyond the scope of this paper as we wish to keep the math to a minimum in the main body of the paper. However, a complete step-by-step derivation and proof of all the results presented is provided in the Supplementary Material as an Appendix.

>>>>

\section{Pruning Neurons to Shrink Neural Networks}\label{sec2}
As discussed in Section \ref{sec1} our aim is to leverage the highly non-uniform distribution of the learning representation in pre-trained neural networks to eliminate redundant neurons, without focusing on individual weight parameters. Taking this approach enables us to remove all the weights (incoming and outgoing) associated with a non-contributing neuron at once. We would like to note here that in an ideal scenario, based on the neuron interdependency theory put forward by \cite{mozer1989skeletonization}, one would evaluate all possible combinations of neurons to remove (one at a time, two at a time, three at a time and so forth) to find the optimal subset of neurons to keep. This is computationally unacceptable, and so we will only focus on removing one neuron at a time and explore more ``greedy'' algorithms to do this in a more efficient manner.

The general approach taken to prune an optimally trained neural network here is to create a ranked list of all the neurons in the network based off of one of the 3 proposed ranking criteria: a brute force approximation, a linear approximation and a quadratic approximation of the neuron's impact on the output of the network. We then test the effects of removing neurons on the accuracy and error of the network. All the algorithms and methods presented here are easily parallelizable as well.

One last thing to note here before moving forward is that the methods discussed in this section involve some non-trivial derivations which are beyond the scope of this paper. We are more focused on analyzing the implications of these methods on our understanding of neural network learning representations. However, a complete step-by-step derivation and proof of all the results presented is provided in the Supplementary Material as an Appendix.
